---
title: "One simulated Beta curve"
author: "Pedro Delicado and Nihan Acar-Denizli"
date: "2026-03-28"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
do_pdf <- FALSE
set.seed(12321)
```

## Defining the curves $\alpha(t)$ and $\beta(t)$

```{r}
# sample size
m <- 1001 # 501 # 201
t <- seq(0,1,length=m)

Cte <- .75
#Cte <- 1
a0 <- 2.5
a1 <- -5
a2 <- 5
#alpha <- Cte*exp((a0+a1*t))
alpha <- exp(Cte*(a0+a1*t+a2*(t-.5)^2))

b0 <- 1.5
b1 <- 5
b2 <- -5
#beta <- Cte*exp((b0+b1*t))
beta <- exp(Cte*(b0+b1*t+b2*t^2))

plot(t,log(alpha), type="l", 
     lwd=2, lty=1,
     ylim=range(c(log(alpha),log(beta))),
     ylab=c("log(alpha), log(beta)"),
     main=c("log(alpha(t)) ___    log(beta(t)) _ _"))
lines(t, log(beta), lwd=2, lty=2)
```


```{r}
plot(t,alpha, type="l", lwd=2, lty=1,
     ylim=range(c(alpha, beta)),
     ylab=expression(paste(alpha(t),",  ",beta(t))),
     main=expression(paste(alpha(t)," ___    ",    beta(t)," _  _")))
lines(t, beta, lwd=2, lty=2)
```


```{r}
mu <- alpha/(alpha+beta)
sigma2 <- alpha*beta/((alpha+beta+1)*(alpha+beta)^2)

q025 <- qbeta(.025,alpha,beta)
q500 <- qbeta(.5,alpha,beta)
q975 <- qbeta(.975,alpha,beta)

plot(t,mu, ylim=c(0,1))
plot(t,sqrt(sigma2))
```

## Random data generation 

```{r}
Y <- rbeta(m,shape1 = alpha, shape2 = beta)

plot(t,Y, ylim=c(0,1), pch=19)
lines(t,mu,lwd=2)
lines(t,mu+qnorm(.9)*sqrt(sigma2),lwd=2,col=2,lty=2)
lines(t,mu+qnorm(.1)*sqrt(sigma2),lwd=2,col=2,lty=2)
```


```{r}
plot(t,Y, ylim=c(0,1), 
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     pch=19, cex=.5, col="darkgreen")
lines(t,mu,lwd=2)
lines(t,q500,lwd=3,col=2,lty=2)
lines(t,q025,lwd=3,col=4,lty=4)
lines(t,q975,lwd=3,col=4,lty=4)

legend("topright",
       c("Mean", "Median", "Q_0.025, Q_0.975"),
       col=c(1,2,4), lty=c(1,2,4), lwd=c(2,3,3),
       bty="n", cex=.5, seg.len = 3.5)
```

**Validation set.**

```{r}
#set.seed(56473)
# sample size for the validation set
m_val <- 1001 #2001
t_val <- seq(0, 1, length=m_val) # sort(runif(m_val))

alpha_tval <- exp(Cte*(a0+a1*t_val + a2*(t_val - .5)^2))
beta_tval <-  exp(Cte*(b0+b1*t_val + b2*t_val^2))

Y_val <- rbeta(m_val, shape1 = alpha_tval, shape2 = beta_tval)

plot(t_val,Y_val, ylim=c(0,1), pch=19)
```


## Estimation

### Reading functions for local maximum likelihood Beta estimation

```{r}
source("generic_functions.R")
source("local_linear_functions.R")
source("local_constant_functions.R")
```



### Estimation: Local constant likelihood

```{r}
# Sample of size m=1001
# t <- t
y <- Y
# size of the grid 'newt'
mt <- 101
newt <- seq(0,1,length=mt)

mv <- FALSE
myBeta <- FALSE
h <- .06 # .075  # (.06 is h_star, found later)

# local constant: 2 parameters


loclikBeta_out_lc_1 <- loc_lik_Beta_2(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv, method="Nelder-Mead")
loclikBeta_out_lc_2 <- loc_lik_Beta_2(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv, method="BFGS", gr=grad_local_log_lik_2)
loclikBeta_out_lc_3 <- loc_lik_Beta_nlm_2(t=t, y=y, h=h, newt=newt, myBeta=myBeta, mv=mv) 

loclikBeta_out <- loclikBeta_out_lc_3

if (mv){
  npar <- dim(loclikBeta_out$ab)[2] # 2 or 4
  par1 <- loclikBeta_out$ab[,1] 
#  par1 <- smooth.spline(newt,par1,df=15)$y
  par2 <- loclikBeta_out$ab[,1+npar/2]
#  par2 <- smooth.spline(newt,par2,df=15)$y
  mu  <- exp(par1)/(1+exp(par1))
  var <- exp(par2)
  aux_denom <- mu*(1-mu)/var - 1
  alphat <- mu*aux_denom   # alpha
  betat <- aux_denom - alphat # beta
  mt <- mu
  st <- sqrt(var)
}else{
  deltat <- loclikBeta_out$deltat 
#  deltat <-  smooth.spline(newt,deltat,df=15)$y
  etat <- loclikBeta_out$etat
#  etat <-  smooth.spline(newt,etat,df=15)$y
  alphat <- exp(deltat) 
  betat <- exp(etat)
  mt <- alphat/(alphat+betat)
  st <- sqrt(mt*(1-mt)/(alphat+betat+1))
}

q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)
```


```{r}
plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen")
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)
```


### Estimation: Local linear likelihood

```{r}
# Sample of size m=1001
# t <- t
y <- Y
# newt <- seq(0,1,length=101) 
mv <- FALSE
myBeta <- FALSE
h <- .12 # .1  (.12 is h_star, found later) 

# local linear: 4 parameters

loclikBeta_out_ll_1 <- loc_lik_Beta(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv)
loclikBeta_out_ll_2 <- loc_lik_Beta(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv, method="BFGS", gr=grad_local_log_lik)
loclikBeta_out_ll_3  <- loc_lik_Beta_nlm(t=t, y=y, h=h, newt=newt, myBeta=myBeta, mv=mv)

loclikBeta_out  <- loclikBeta_out_ll_2

if (mv){
  npar <- dim(loclikBeta_out$ab)[2] # 2 or 4
  par1 <- loclikBeta_out$ab[,1] 
#  par1 <- smooth.spline(newt,par1,df=15)$y
  par2 <- loclikBeta_out$ab[,1+npar/2]
#  par2 <- smooth.spline(newt,par2,df=15)$y
  mu  <- exp(par1)/(1+exp(par1))
  var <- exp(par2)
  aux_denom <- mu*(1-mu)/var - 1
  alphat <- mu*aux_denom   # alpha
  betat <- aux_denom - alphat # beta
  mt <- mu
  st <- sqrt(var)
}else{
  deltat <- loclikBeta_out$deltat 
#  deltat <-  smooth.spline(newt,deltat,df=15)$y
  etat <- loclikBeta_out$etat
#  etat <-  smooth.spline(newt,etat,df=15)$y
  alphat <- exp(deltat) 
  betat <- exp(etat)
  mt <- alphat/(alphat+betat)
  st <- sqrt(mt*(1-mt)/(alphat+betat+1))
}

q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)
```


```{r}
plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen")
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)
```

### Comparing 6 different estimators

#### Estimation performance

```{r,fig.width=7.2,fig.height=3.6}
if (do_pdf) pdf(file="One_Sim_Beta_m_201_6_estims.pdf",
                width=7.2,height=3.6)
op <-par(mfrow=c(1,2))
plot(t,alpha,type="l",lwd=2, ylab=expression(alpha(t)))
lines(newt, exp(loclikBeta_out_lc_1$deltat),lwd=2, col=5, lty=2)
lines(newt, exp(loclikBeta_out_lc_2$deltat),lwd=2, col=3, lty=2)
lines(newt, exp(loclikBeta_out_lc_3$deltat),lwd=2, col=2, lty=2)
lines(newt, exp(loclikBeta_out_ll_1$deltat),lwd=2, col=7, lty=4)
lines(newt, exp(loclikBeta_out_ll_2$deltat),lwd=2, col=6, lty=4)
lines(newt, exp(loclikBeta_out_ll_3$deltat),lwd=2, col=4, lty=4)
legend("topright",
       c(expression(paste("True ", alpha,"(t)")),
         "3 local constant estims", 
         "3 local linear estims"),
       col=c(1,2,4),lwd=2,
       lty=c(1,2,4),
       bty="n", cex=.75, seg.len = 3.5)

plot(t,beta,type="l",lwd=2, ylab=expression(beta(t)), ylim=range(alpha))
lines(newt, exp(loclikBeta_out_lc_1$eta),lwd=2, col=5, lty=2)
lines(newt, exp(loclikBeta_out_lc_2$eta),lwd=2, col=3, lty=2)
lines(newt, exp(loclikBeta_out_lc_3$eta),lwd=2, col=2, lty=2)
lines(newt, exp(loclikBeta_out_ll_1$eta),lwd=2, col=7, lty=4)
lines(newt, exp(loclikBeta_out_ll_2$eta),lwd=2, col=6, lty=4)
lines(newt, exp(loclikBeta_out_ll_3$eta),lwd=2, col=4, lty=4)
legend("topright",
       c(expression(paste("True ", beta,"(t)")),
         "3 local constant estims", 
         "3 local linear estims"),
       col=c(1,2,4),lwd=2,
       lty=c(1,2,4),
       bty="n", cex=.75, seg.len = 3.5)
par(op)

if (do_pdf) dev.off()
```

#### Computing time

```{r}
library(microbenchmark)
```


```{r, eval=FALSE, echo=FALSE}
mb_times <- microbenchmark(
  lc_Nelder_Mead=loc_lik_Beta_2(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv),
  lc_BFGS=loc_lik_Beta_2(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv, method="BFGS", gr=grad_local_log_lik_2),
  lc_NLM=loc_lik_Beta_nlm_2(t=t, y=y, h=h, newt=newt, myBeta=myBeta, mv=mv),
  ll_Nelder_Mead=loc_lik_Beta(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv),
  ll_BFGS=loc_lik_Beta(t=t,y=y, h=h, newt=newt, myBeta=myBeta, mv=mv, method="BFGS", gr=grad_local_log_lik),
  ll_NLM=loc_lik_Beta_nlm(t=t, y=y, h=h, newt=newt, myBeta=myBeta, mv=mv),
  times=10
)
```


```{r, fig.width=7,fig.height=5, eval=FALSE, echo=FALSE}
#if (do_pdf) pdf(file="box_plot_micro_bench_m_201.pdf", width=7, height=5)
if (do_pdf) pdf(file="box_plot_micro_bench_m_1001.pdf", width=7, height=5)
summary(mb_times)
# m=201:
##
# First submission: October 2025
#Unit: milliseconds
#             expr       min        lq       mean     median        uq       max neval
# 1 lc_Nelder_Mead  218.5858  239.9247  250.22380  243.63100  250.3775  298.1790    10
# 2        lc_BFGS   79.7568   82.7506  106.47682   88.07505  104.9944  200.5937    10
# 3         lc_NLM   44.3620   45.8548   55.94153   50.59635   55.1942   94.8627    10
# 4 ll_Nelder_Mead 1854.0138 1939.2801 2140.04123 2162.27490 2204.3362 2606.0346    10
# 5        ll_BFGS  433.7535  469.0023  517.67281  504.63940  527.2358  655.7318    10
# 6         ll_NLM  446.1252  454.3509  512.62236  492.01160  538.4956  732.9094    10
#
##
# Second submission: March 2026
# Unit: milliseconds
#            expr       min        lq       mean     median        uq       max neval   cld
#  lc_Nelder_Mead  169.1762  172.9005  177.91147  177.96825  181.6806  188.5828    10 a    
#         lc_BFGS   66.8573   68.3761   74.23814   74.03155   80.3090   84.3417    10  b   
#          lc_NLM   38.9405   39.4720   43.96864   41.38160   50.6137   51.4918    10  b   
#  ll_Nelder_Mead 1476.0505 1491.7396 1540.29687 1546.53740 1568.8772 1643.7574    10   c  
#         ll_BFGS  308.2743  315.0886  327.27678  320.09910  339.0796  366.4639    10    d 
#          ll_NLM  359.6628  374.0655  382.20985  378.43725  384.2932  424.1071    10     e
####
# m=1001:
##
# First submission: October 2025
# Unit: milliseconds
#            expr       min        lq      mean    median        uq        max neval
#  lc_Nelder_Mead 1091.6772 1169.9430 1212.5945 1226.2528 1255.3913  1300.0997    10
#         lc_BFGS  528.2959  542.0339  568.9238  569.8263  586.5533   617.1882    10
#          lc_NLM  238.7890  252.3196  262.7388  262.4054  275.5777   281.4746    10
#  ll_Nelder_Mead 9414.1943 9582.6263 9750.1471 9741.9744 9920.1234 10080.0480    10
#         ll_BFGS 1630.2858 1656.0706 1700.2701 1699.5673 1739.3714  1792.3576    10
#          ll_NLM 2252.2240 2286.9975 2331.0192 2315.5427 2368.9760  2494.0073    10
##
# Second submission: March 2026
# Unit: milliseconds
#            expr       min        lq      mean    median        uq       max neval    cld
#  lc_Nelder_Mead  746.8826  752.0027  769.8930  759.9375  769.5706  826.3411    10 a     
#         lc_BFGS  381.0127  387.1707  395.0858  395.6788  400.2642  410.3662    10  b    
#          lc_NLM  151.6345  154.5403  157.6115  155.2358  157.4352  178.4908    10   c   
#  ll_Nelder_Mead 6325.5503 6376.4893 6501.9413 6414.5394 6481.9302 7201.1633    10    d  
#         ll_BFGS 1114.2651 1126.8117 1174.6135 1152.8555 1217.1820 1318.5205    10     e 
#          ll_NLM 1675.0760 1686.8989 1752.7208 1756.0758 1781.8698 1881.4473    10      f
#
boxplot(mb_times, las=2, xlab="", cex.axis=.6)
if (do_pdf) dev.off()
```

```{r, echo=FALSE, eval=FALSE}
if (m==201){
  save.image(file="One_Simulated_Beta_Curve_m_201.Rdata")
  # save.image(file="One_Simulated_Beta_Curve_m_201_March_2026_Portatil_Pedro.Rdata")
}
if (m==1001){
  save.image(file="One_Simulated_Beta_Curve_m_1001.Rdata")
  # save.image(file="One_Simulated_Beta_Curve_m_1001_March_2026_Portatil_Pedro.Rdata")
}
```


**Conclusions: (Second submission, March 2026)**

The 3 optimization methods produce essentially the same estimation results, with different computation times:
  
  - In local constant, NLM is the fastest method, followed by BFGS and them Nelder-Mead
  - In local linear BFGS is the fastest method, followed by NLM and them Nelder-Mead
  - For $m=201$, Local constant NLM is around 7.7 times faster than local linear BFGS.
  - For $m=1001$, Local constant NLM is around 7.4 times faster than local linear BFGS.
  - For $m=1001$ the computation times are between 3.6 and 5.3 times those of $m=201$ (time less than proportional to sample size $m$).
  
So we decide: 
*"we will use nlm for local constant likelihood estimation, and optim with BFGS for local linear cases."*


### Bandwidth choice 

#### Local constant Beta: looCV, approximated looCV, and k-fold CV for one curve

```{r} 
Kh <- 15
vh_lc_201  <- seq(.03,.07, length=Kh) # range(h_val_set) 100 simul
vh_lc_501  <- seq(.02,.055, length=Kh) # 
vh_lc_1001 <- seq(.014,.045, length=Kh)# 
vh_ll_201  <- seq(.058,.2, length=Kh)  # 
vh_ll_501  <- seq(.032,.175, length=Kh) # 
vh_ll_1001 <- seq(.032,.16, length=Kh)# 


if (m==201) {
  vh_lc <- vh_lc_201
  vh_ll <- vh_ll_201
}
if (m==501) {
  vh_lc <- vh_lc_501
  vh_ll <- vh_ll_501
}
if (m==1001) {
  vh_lc <- vh_lc_1001
  vh_ll <- vh_ll_1001
}
```


```{r} 
system.time(
loo_1sim <- looCV_loclikBeta_2(t, y,
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE)
)
       # m=201:
       # usuari       sistema transcorregut 
       #   0.66          0.00          0.75
#
       # m=1001:
       # usuari       sistema transcorregut 
       #  10.60          0.09         11.19 

h_star <- h_loo <- loo_1sim$vh[which.max(loo_1sim$loglik)] 

plot(loo_1sim$vh, loo_1sim$loglik/m)
abline(v=h_loo,col=2)
```

```{r} 
system.time(
infl_1sim <- infl_looCV_loclikBeta_2nd_2(t, y, newt = newt,
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE)
)
       # m=201
       # usuari       sistema transcorregut 
       #   2.36          0.19          2.70
#
       # m=1001
       # usuari       sistema transcorregut 
       # 170.31         22.12        199.03
         
h_star <- h_infl <- infl_1sim$vh[which.max(infl_1sim$loologlik)]

plot(infl_1sim$vh, infl_1sim$loologlik/m)
abline(v=h_infl, col=2)
```

```{r} 
system.time(
kfCV_1sim <- k_fold_CV_loclikBeta_2(t, y,
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE)
)
       # m=201:
       # usuari       sistema transcorregut 
       #   0.62          0.00          0.69
#
       # m=1001:
       # usuari       sistema transcorregut 
       #   8.30          0.10          8.73 

h_star_lc <- h_star <- h_kfCV <- kfCV_1sim$vh[which.max(kfCV_1sim$loglik)]

plot(kfCV_1sim$vh, kfCV_1sim$loglik/m)
abline(v=h_kfCV,col=2)
```

```{r log_lik_validation_set}
Kh <- length(vh_lc)
mean_log_lik_val_set_lc <- numeric(Kh)
for (k in 1:Kh){
  mean_log_lik_val_set_lc[k] <- 
    log_lik_Beta_val_set(tval=t_val, yval=Y_val, 
#    log_lik_Beta_val_set(tval=t, yval=Y, # training sample
                         newt=newt, 
                         deltant=infl_1sim$deltant[k,], 
                         etant=infl_1sim$etant[k,])
}
h_star_lc <- h_val_set <- vh_lc[which.max(mean_log_lik_val_set_lc)]
plot(vh_lc, mean_log_lik_val_set_lc)
abline(v=h_val_set,col=2)
```

```{r, fig.width=6, fig.asp=1}
rg_log_lik <- range(c(mean_log_lik_val_set_lc, loo_1sim$loglik/m,
                infl_1sim$loologlik/m, kfCV_1sim$loglik/m))
dh <- diff(range(vh_lc))/201
plot(vh_lc, mean_log_lik_val_set_lc, type="b", col="black", lwd=2, 
     ylim=rg_log_lik, xlab="h", 
     ylab="log-likelihood", 
     main="Estimates of the log-likelihood mean (local constant)")
abline(v=h_val_set,col="black", lty=2)
lines(vh_lc, loo_1sim$loglik/m, type="b", col="red", lwd=2)
abline(v=h_loo-dh/2, col="red", lty=2)
lines(vh_lc, infl_1sim$loologlik/m, type="b", col="darkgreen", lwd=2)
abline(v=h_infl+dh, col="darkgreen", lty=2)
lines(vh_lc, kfCV_1sim$loglik/m, type="b", col="blue", lwd=2)
abline(v=h_kfCV+dh/2, col="blue", lty=2)
legend("topright",
       c("Valid.set", "loo-CV", "approx.loo", "k-fold-CV"),
       col=c("black", "red", "darkgreen", "blue"), lty=1, lwd=2,
       bty="n", cex=.8, seg.len = 3.5)
```


```{r}
loclikBeta_out_lc <- loclikBeta_out <- 
  loc_lik_Beta_nlm_2(t=t, y=y, h=h_star_lc, 
                     newt=newt, myBeta=myBeta, mv=mv)
if (mv){
  npar <- dim(loclikBeta_out$ab)[2] # 2 or 4
  par1 <- loclikBeta_out$ab[,1] 
#  par1 <- smooth.spline(newt,par1,df=15)$y
  par2 <- loclikBeta_out$ab[,1+npar/2]
#  par2 <- smooth.spline(newt,par2,df=15)$y
  mu  <- exp(par1)/(1+exp(par1))
  var <- exp(par2)
  aux_denom <- mu*(1-mu)/var - 1
  alphat <- mu*aux_denom   # alpha
  betat <- aux_denom - alphat # beta
  mt <- mu
  st <- sqrt(var)
}else{
  deltat <- loclikBeta_out$deltat 
#  deltat <-  smooth.spline(newt,deltat,df=15)$y
  etat <- loclikBeta_out$etat
#  etat <-  smooth.spline(newt,etat,df=15)$y
  alphat <- exp(deltat) 
  betat <- exp(etat)
  mt <- alphat/(alphat+betat)
  st <- sqrt(mt*(1-mt)/(alphat+betat+1))
}

q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("h_star=",h_star_lc))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)
```


#### Local linear Beta: looCV, approximated looCV, and k-fold CV for one curve

```{r} 
system.time(
loo_1sim <- looCV_loclikBeta(t, y,
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE)
)
       # m=201:
       # usuari       sistema transcorregut 
       #   4.78          0.04          5.14 
#
       # m=1001:
       # usuari       sistema transcorregut 
       #  99.44          1.14        104.05 

h_star <- h_loo <- loo_1sim$vh[which.max(loo_1sim$loglik)]

plot(loo_1sim$vh, loo_1sim$loglik/m)
abline(v=h_loo,col=2)
```

```{r} 
system.time(
infl_1sim <- infl_looCV_loclikBeta_2nd(t, y, newt = newt,
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE)
)
       # m=201:
       # usuari       sistema transcorregut 
       #   7.48          0.13          8.08
#
       # m=1001:
       # usuari       sistema transcorregut 
       # 305.09         28.85        341.17
         
h_star <- h_infl <- infl_1sim$vh[which.max(infl_1sim$loologlik)]

plot(infl_1sim$vh, infl_1sim$loologlik/m)
abline(v=h_infl,col=2)
```

```{r} 
system.time(
kfCV_1sim <- k_fold_CV_loclikBeta(t, y,
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE)
)
       # m=201:
       # usuari       sistema transcorregut 
       #   3.91          0.03          4.16
#
       # m=1001:
       # usuari       sistema transcorregut 
       #  77.77          0.90         80.88
         
h_star_ll <- h_star <- h_kfCV <- kfCV_1sim$vh[which.max(kfCV_1sim$loglik)]

plot(kfCV_1sim$vh, kfCV_1sim$loglik/m)
abline(v=h_kfCV,col=2)
```

```{r log_lik_validation_set_ll}
Kh <- length(vh_ll)
mean_log_lik_val_set_ll <- numeric(Kh)
for (k in 1:Kh){
  mean_log_lik_val_set_ll[k] <- 
    log_lik_Beta_val_set(tval=t_val, yval=Y_val, 
#    log_lik_Beta_val_set(tval=t, yval=Y, # training sample
                         newt=newt, 
                         deltant=infl_1sim$deltant[k,], 
                         etant=infl_1sim$etant[k,])
}
h_star_ll <- h_val_set <- vh_ll[which.max(mean_log_lik_val_set_ll)]
plot(vh_ll, mean_log_lik_val_set_ll)
abline(v=h_val_set,col=2)
```

```{r, fig.width=6, fig.asp=1}
rg_log_lik <- range(c(mean_log_lik_val_set_ll, loo_1sim$loglik/m,
                infl_1sim$loologlik/m, kfCV_1sim$loglik/m))
dh <- diff(range(vh_ll))/201
plot(vh_ll, mean_log_lik_val_set_ll, type="b", col="black", lwd=2, 
     ylim=rg_log_lik, xlab="h", 
     ylab="log-likelihood", 
     main="Estimates of the log-likelihood mean (local linear)")
abline(v=h_val_set,col="black", lty=2)
lines(vh_ll, loo_1sim$loglik/m, type="b", col="red", lwd=2)
abline(v=h_loo-dh/2, col="red", lty=2)
lines(vh_ll, infl_1sim$loologlik/m, type="b", col="darkgreen", lwd=2)
abline(v=h_infl+dh, col="darkgreen", lty=2)
lines(vh_ll, kfCV_1sim$loglik/m, type="b", col="blue", lwd=2)
abline(v=h_kfCV+dh/2, col="blue", lty=2)
legend("topright",
       c("Valid.set", "loo-CV", "approx.loo", "k-fold-CV"),
       col=c("black", "red", "darkgreen", "blue"), lty=1, lwd=2,
       bty="n", cex=.8, seg.len = 3.5)
```

```{r}
loclikBeta_out_ll <- loclikBeta_out <- 
  loc_lik_Beta(t=t, y=y, h=h_star_ll, method = "BFGS",  
                     newt=newt, myBeta=myBeta, mv=mv)
if (mv){
  npar <- dim(loclikBeta_out$ab)[2] # 2 or 4
  par1 <- loclikBeta_out$ab[,1] 
#  par1 <- smooth.spline(newt,par1,df=15)$y
  par2 <- loclikBeta_out$ab[,1+npar/2]
#  par2 <- smooth.spline(newt,par2,df=15)$y
  mu  <- exp(par1)/(1+exp(par1))
  var <- exp(par2)
  aux_denom <- mu*(1-mu)/var - 1
  alphat <- mu*aux_denom   # alpha
  betat <- aux_denom - alphat # beta
  mt <- mu
  st <- sqrt(var)
}else{
  deltat <- loclikBeta_out$deltat 
#  deltat <-  smooth.spline(newt,deltat,df=15)$y
  etat <- loclikBeta_out$etat
#  etat <-  smooth.spline(newt,etat,df=15)$y
  alphat <- exp(deltat) 
  betat <- exp(etat)
  mt <- alphat/(alphat+betat)
  st <- sqrt(mt*(1-mt)/(alphat+betat+1))
}

q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("h_star=",h_star_ll))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)
```


#### Microbenchmarking the bandwidth choice methods

```{r, eval=FALSE, echo=FALSE}
library(microbenchmark)
band_choice_times <- microbenchmark(
  lc_loo= looCV_loclikBeta_2(t, y,
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE),
  lc_infl_newt= infl_looCV_loclikBeta_2nd_2(t, y, newt = newt, 
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE),
  lc_infl_t= infl_looCV_loclikBeta_2nd_2(t, y, newt = t, 
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE),
  lc_kfold= k_fold_CV_loclikBeta_2(t, y,
                 vh=vh_lc,
                 method="nlm", gr=grad_local_log_lik_2, 
                 myBeta=FALSE, mv=FALSE),
  ll_loo= looCV_loclikBeta(t, y,
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE),
  ll_infl_newt= infl_looCV_loclikBeta_2nd(t, y, newt = newt, 
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE),
  ll_infl_t= infl_looCV_loclikBeta_2nd(t, y, newt = t, 
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE),
  ll_kfold= k_fold_CV_loclikBeta(t, y,
                 vh=vh_ll,
                 method="BFGS", gr=grad_local_log_lik, 
                 myBeta=FALSE, mv=FALSE),
  times=1#10
)
```


```{r, fig.width=7,fig.height=5, eval=FALSE, echo=FALSE}
if (do_pdf) pdf(file="box_plot_band_choice_micro_bench_m_201.pdf", width=7, height=5)
summary(band_choice_times)

# First submission: October 2025
# m=201:
#Unit: milliseconds
#       expr        min         lq       mean     median         uq        max neval
# 1   lc_loo   525.2357   617.7481   677.8130   665.0854   756.8433   816.5764    10
# 2  lc_infl  1752.8510  1933.6777  2227.0891  2241.4037  2509.1642  2690.1614    10
# 3 lc_kfold   445.4487   468.7593   507.7047   500.1790   544.4316   600.9868    10
# 4   ll_loo  8341.0759  8399.1707  9798.2985  9627.3201 10992.4309 11633.0110    10
# 5  ll_infl 10814.6960 11049.5124 12788.4395 11610.1223 15000.6917 15525.6238    10
# 6 ll_kfold  8295.9996  8451.5038  9517.7376  9643.1922 10283.9070 11475.2595    10

# First submission: October 2025
# m=1001:
# Unit: seconds
#      expr        min         lq       mean     median         uq       max neval
#    lc_loo  11.556392  11.760065  12.215925  12.111247  12.630892  13.23503    10
#   lc_infl 188.869230 190.804120 194.360366 192.836068 197.043645 203.33102    10
#  lc_kfold   9.019851   9.391309   9.623293   9.544927   9.796315  10.45609    10
#    ll_loo 135.517630 140.330806 151.830124 146.554265 162.021360 185.84823    10
#   ll_infl 343.550949 343.669805 349.527011 345.984666 356.156404 360.80805    10
#  ll_kfold 111.513907 113.497857 117.563927 116.033302 117.839438 137.13751    10

# Second submission: March 2026
# m=201, with improvements indicated by a referee (l*_infl_t) and additionally using newt (l*_infl_newt)
#Unit: milliseconds
#           expr       min        lq      mean    median        uq       max neval    cld
# 1       lc_loo  539.7746  572.2644  614.5809  593.5570  632.3777  784.4259    10 ab    
# 2 lc_infl_newt  388.3701  400.3374  430.4969  416.0108  444.9771  538.4935    10 a     
# 3    lc_infl_t  648.7639  657.1474  694.9789  681.4574  706.7859  840.4391    10  b    
# 4     lc_kfold  468.3939  480.8862  515.2807  497.4156  509.5903  683.6932    10 ab    
# 5       ll_loo 3700.1130 4059.0655 4128.8279 4110.6838 4220.1044 4401.2842    10   c   
# 6 ll_infl_newt 2490.9220 2548.9307 2673.5424 2683.3203 2796.9169 2858.3282    10    d  
# 7    ll_infl_t 4317.4134 4451.4858 4706.9631 4655.3300 4894.7835 5460.4156    10     e 
# 8     ll_kfold 3224.7753 3357.2116 3559.4877 3417.5647 3784.1172 4214.0067    10      f

# Second submission: March 2026
# m=1001, with improvements indicated by a referee (l*_infl_t) and additionally using newt (l*_infl_newt)
# Unit: seconds
#           expr       min        lq       mean    median        uq        max neval cld
# 1       lc_loo  9.183106  9.229596   9.346369  9.325855  9.429029   9.582135    10  a 
# 2 lc_infl_newt  2.245036  2.269341   2.365959  2.370270  2.424894   2.552128    10  a 
# 3    lc_infl_t 10.323495 10.489311  13.865372 10.840471 11.762904  38.266102    10  a 
# 4     lc_kfold  7.406629  7.537675   8.183831  7.642478  9.092389  10.011366    10  a 
# 5       ll_loo 68.068805 68.509056  73.004088 69.572200 71.251179  98.533045    10   b
# 6 ll_infl_newt 14.853825 15.072270  15.312813 15.117562 15.273266  16.970369    10  a 
# 7    ll_infl_t 76.704342 78.074092 103.118566 79.959147 96.455950 277.213510    10   b
# 8     ll_kfold 55.902835 56.007436  71.802726 56.656960 57.865140 199.249612    10   b

boxplot(band_choice_times, las=2, xlab="", cex.axis=.6)
if (do_pdf) dev.off()
```




## 2x2 plots in 1 page 

```{r, fig.width=7.2, fig.height=7}
if (do_pdf) pdf("2x2_One_Beta_Simul.pdf", 
                width = 7.2, height = 7)
op<-par(mfrow=c(2,2))

# plot 1
plot(t,alpha, type="l", lwd=2, lty=1,
     ylim=range(c(alpha, beta)),
     ylab=expression(paste(alpha(t),",  ",beta(t))),
     main="True Beta functional parameters")
lines(t, beta, lwd=2, lty=2)
legend("topright",
       c(expression(alpha(t)),
         expression(beta(t))), 
       col=1, lty=c(1,2), lwd=2,
       bty="n", cex=1, seg.len = 3.5)
# plot 2
plot(t,Y, ylim=c(0,1), 
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     pch=19, cex=.5, col="darkgreen",
     main="True distributions and random data")
lines(t,mu,lwd=2)
lines(t,q500,lwd=3,col=2,lty=2)
lines(t,q025,lwd=3,col=4,lty=4)
lines(t,q975,lwd=3,col=4,lty=4)

legend("topright",
       c("Mean(t)", "Median(t)", 
         expression(Q[.025](t)),
         expression(Q[.975](t))),
       col=c(1,2,4), lty=c(1,2,4), lwd=c(2,3,3),
       bty="n", cex=1, seg.len = 3.5)

# plot 3
loclikBeta_out <- loclikBeta_out_lc
deltat <- loclikBeta_out$deltat 
etat <- loclikBeta_out$etat
alphat <- exp(deltat) 
betat <- exp(etat)
mt <- alphat/(alphat+betat)
st <- sqrt(mt*(1-mt)/(alphat+betat+1))
q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("Local constant estimation, h*=",h_star_lc))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)

# plot 4
loclikBeta_out <- loclikBeta_out_ll
deltat <- loclikBeta_out$deltat 
etat <- loclikBeta_out$etat
alphat <- exp(deltat) 
betat <- exp(etat)
mt <- alphat/(alphat+betat)
st <- sqrt(mt*(1-mt)/(alphat+betat+1))
q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("Local linear estimation, h*=",h_star_ll))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)

par(op)
if (do_pdf) dev.off()
```




## 3x2 plots in 1 page. Version A

```{r, fig.width=7.2, fig.height=10.5}
if (do_pdf) pdf("3x2_One_Beta_Simul_A.pdf", 
                width = 7.2, height = 7)
op<-par(mfrow=c(3,2))

# plot 1
plot(t,alpha, type="l", lwd=2, lty=1,
     ylim=range(c(alpha, beta)),
     ylab=expression(paste(alpha(t),",  ",beta(t))),
     main="True Beta functional parameters")
lines(t, beta, lwd=2, lty=2)
legend("topright",
       c(expression(alpha(t)),
         expression(beta(t))), 
       col=1, lty=c(1,2), lwd=2,
       bty="n", cex=1, seg.len = 3.5)
# plot 2
plot(t,Y, ylim=c(0,1), 
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     pch=19, cex=.5, col="darkgreen",
     main="True distributions and random data")
lines(t,mu,lwd=2)
lines(t,q500,lwd=3,col=2,lty=2)
lines(t,q025,lwd=3,col=4,lty=4)
lines(t,q975,lwd=3,col=4,lty=4)

legend("topright",
       c("Mean(t)", "Median(t)", 
         expression(Q[.025](t)),
         expression(Q[.975](t))),
       col=c(1,2,4), lty=c(1,2,4), lwd=c(2,3,3),
       bty="n", cex=1, seg.len = 3.5)


# Plot 3 

plot(t,alpha,type="l",lwd=2, ylab=expression(alpha(t)))
lines(newt, exp(loclikBeta_out_lc_1$deltat),lwd=2, col=5, lty=2)
lines(newt, exp(loclikBeta_out_lc_2$deltat),lwd=2, col=3, lty=2)
lines(newt, exp(loclikBeta_out_lc_3$deltat),lwd=2, col=2, lty=2)
lines(newt, exp(loclikBeta_out_ll_1$deltat),lwd=2, col=7, lty=4)
lines(newt, exp(loclikBeta_out_ll_2$deltat),lwd=2, col=6, lty=4)
lines(newt, exp(loclikBeta_out_ll_3$deltat),lwd=2, col=4, lty=4)
legend("topright",
       c(expression(paste("True ", alpha,"(t)")),
         "3 local constant estims", 
         "3 local linear estims"),
       col=c(1,2,4),lwd=2,
       lty=c(1,2,4),
       bty="n", cex=.75, seg.len = 3.5)


# plot 4

plot(t,beta,type="l",lwd=2, ylab=expression(beta(t)), ylim=range(alpha))
lines(newt, exp(loclikBeta_out_lc_1$eta),lwd=2, col=5, lty=2)
lines(newt, exp(loclikBeta_out_lc_2$eta),lwd=2, col=3, lty=2)
lines(newt, exp(loclikBeta_out_lc_3$eta),lwd=2, col=2, lty=2)
lines(newt, exp(loclikBeta_out_ll_1$eta),lwd=2, col=7, lty=4)
lines(newt, exp(loclikBeta_out_ll_2$eta),lwd=2, col=6, lty=4)
lines(newt, exp(loclikBeta_out_ll_3$eta),lwd=2, col=4, lty=4)
legend("topright",
       c(expression(paste("True ", beta,"(t)")),
         "3 local constant estims", 
         "3 local linear estims"),
       col=c(1,2,4),lwd=2,
       lty=c(1,2,4),
       bty="n", cex=.75, seg.len = 3.5)

# plot 5
loclikBeta_out <- loclikBeta_out_lc
deltat <- loclikBeta_out$deltat 
etat <- loclikBeta_out$etat
alphat <- exp(deltat) 
betat <- exp(etat)
mt <- alphat/(alphat+betat)
st <- sqrt(mt*(1-mt)/(alphat+betat+1))
q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("Local constant estimation, h*=",h_star_lc))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)

# plot 6
loclikBeta_out <- loclikBeta_out_ll
deltat <- loclikBeta_out$deltat 
etat <- loclikBeta_out$etat
alphat <- exp(deltat) 
betat <- exp(etat)
mt <- alphat/(alphat+betat)
st <- sqrt(mt*(1-mt)/(alphat+betat+1))
q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("Local linear estimation, h*=",h_star_ll))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)

par(op)
if (do_pdf) dev.off()
```


## 3x2 plots in 1 page. Version B

```{r, fig.width=6, fig.height=8.75}
if (do_pdf) pdf("3x2_One_Beta_Simul_B.pdf", 
                width = 6, height = 8.75)
op<-par(mfrow=c(3,2))

# plot 1
plot(t,alpha, type="l", lwd=2, lty=1,
     ylim=range(c(alpha, beta)),
     ylab=expression(paste(alpha(t),",  ",beta(t))),
     main="True Beta functional parameters")
lines(t, beta, lwd=2, lty=2)
legend("topright",
       c(expression(alpha(t)),
         expression(beta(t))), 
       col=1, lty=c(1,2), lwd=2,
       bty="n", cex=1, seg.len = 3.5)
# plot 2
plot(t,Y, ylim=c(0,1), 
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     pch=19, cex=.5, col="darkgreen",
     main="True distributions and random data")
lines(t,mu,lwd=2)
lines(t,q500,lwd=3,col=2,lty=2)
lines(t,q025,lwd=3,col=4,lty=4)
lines(t,q975,lwd=3,col=4,lty=4)

legend("topright",
       c("Mean(t)", "Median(t)", 
         expression(Q[.025](t)),
         expression(Q[.975](t))),
       col=c(1,2,4), lty=c(1,2,4), lwd=c(2,3,3),
       bty="n", cex=1, seg.len = 3.5)


# Plot 3 

plot(t,alpha,type="l",lwd=2, 
     ylab=expression(alpha(t)),
          main=expression(paste("Estimation of ", alpha(t))))
lines(newt, exp(loclikBeta_out_lc_1$deltat),lwd=2, col=5, lty=2)
lines(newt, exp(loclikBeta_out_lc_2$deltat),lwd=2, col=3, lty=2)
lines(newt, exp(loclikBeta_out_lc_3$deltat),lwd=2, col=2, lty=2)
lines(newt, exp(loclikBeta_out_ll_1$deltat),lwd=2, col=7, lty=4)
lines(newt, exp(loclikBeta_out_ll_2$deltat),lwd=2, col=6, lty=4)
lines(newt, exp(loclikBeta_out_ll_3$deltat),lwd=2, col=4, lty=4)
legend("topright",
       c(expression(paste("True ", alpha,"(t)")),
         "3 local constant estims", 
         "3 local linear estims"),
       col=c(1,2,4),lwd=2,
       lty=c(1,2,4),
       bty="n", cex=.75, seg.len = 3.5)


# plot 4
loclikBeta_out <- loclikBeta_out_lc
deltat <- loclikBeta_out$deltat 
etat <- loclikBeta_out$etat
alphat <- exp(deltat) 
betat <- exp(etat)
mt <- alphat/(alphat+betat)
st <- sqrt(mt*(1-mt)/(alphat+betat+1))
q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("Local constant estimation, h*=",h_star_lc))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)

# plot 5

plot(t,beta,type="l",lwd=2, 
     ylab=expression(beta(t)),
     main=expression(paste("Estimation of ", beta(t))),
     ylim=range(alpha))
lines(newt, exp(loclikBeta_out_lc_1$eta),lwd=2, col=5, lty=2)
lines(newt, exp(loclikBeta_out_lc_2$eta),lwd=2, col=3, lty=2)
lines(newt, exp(loclikBeta_out_lc_3$eta),lwd=2, col=2, lty=2)
lines(newt, exp(loclikBeta_out_ll_1$eta),lwd=2, col=7, lty=4)
lines(newt, exp(loclikBeta_out_ll_2$eta),lwd=2, col=6, lty=4)
lines(newt, exp(loclikBeta_out_ll_3$eta),lwd=2, col=4, lty=4)
legend("topright",
       c(expression(paste("True ", beta,"(t)")),
         "3 local constant estims", 
         "3 local linear estims"),
       col=c(1,2,4),lwd=2,
       lty=c(1,2,4),
       bty="n", cex=.75, seg.len = 3.5)

# plot 6
loclikBeta_out <- loclikBeta_out_ll
deltat <- loclikBeta_out$deltat 
etat <- loclikBeta_out$etat
alphat <- exp(deltat) 
betat <- exp(etat)
mt <- alphat/(alphat+betat)
st <- sqrt(mt*(1-mt)/(alphat+betat+1))
q025t <- qbeta(.025,alphat,betat)
q500t <- qbeta(.5,alphat,betat)
q975t <- qbeta(.975,alphat,betat)

plot(t,y, ylim=c(0,1), pch=19, cex=.5,col="darkgreen",
     xlab=expression(t[j]),
     ylab=expression(Y(t[j])),
     main=paste("Local linear estimation, h*=",h_star_ll))
lines(newt,mt,col=1,lwd=2)
lines(newt,q500t,col=2,lty=2,lwd=3)
lines(newt,q025t,col=4,lty=4,lwd=3)
lines(newt,q975t,col=4,lty=4,lwd=3)

par(op)
if (do_pdf) dev.off()
```